Transfer learning for \'echelle spectroscopy with \texttt{blas\'e}

### 2022-03-24 paper/paper1/ms.tex
Add description of exponentially spaced pixel samples

But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel overall, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.  We therefore urge some caution when applying \texttt{blas\'e} to telluric modeling in the most saturated regions.  \authorcomment1{An obvious and good solution is to intialize the model in log flux.  We may revise the scheme to do so.}

@@

\subsection{Augmenting the Stellar Clone Model for $RV$ and $v\sin{i}$ }
Real stars possess two key extrinsic properties.  Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location. The qualifier \emph{extrinsic} distiguishes from stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$, which appear the same from any viewing location, at least for stars with isotropic surfaces.  The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.  We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:

\begin{eqnarray}
    \mathsf{S}_{\rm ext}(\bm{\lambda}_S) = \mathsf{S}_{\rm clone}(\lambda_\mathrm{c} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
\end{eqnarray}

@@

Operationally, the Doppler shift gets applied to the line center positions and not the individual wavelength grid point coordinates, $\bm{\lambda}_S$. This choice yields a key benefit: it makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can easily be sensed through backpropagation by affecting only the line center positions.  The strategy allows relatively low computational and memory cost since it comports with the sparse matrix implementation.

@@

\begin{eqnarray}
    \mathsf{M}_{\rm joint} = \mathsf{S}_{\rm ext}(\bm{\lambda}_S) \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_S)
\end{eqnarray}

>>>>

But overall, the telluric model cloning performs adequately, with \censorbox{X percentage} error-per-pixel overall, and only \censorbox{zz} error-per-pixel when restricted to the atmospheric windows.

@@

\subsection{Augmenting the Stellar Clone Model for $RV$ and $v\sin{i}$ }
Real stars possess two key extrinsic properties.  Rotational broadening $v\sin{i}$ and radial velocity $RV$ depend on the observer's viewing location. We follow \citet{czekala15} by emphasizing the qualifier \emph{extrinsic}, to distiguish between stellar \emph{intrinsic} properties, such as $T_{\mathrm{eff}}, \log{g},\mathrm{and\,} [\mathrm{Fe}/\mathrm{H}]$.  Intrinsic properties appear the same from any viewing location---at least for stars with isotropic surfaces---while extrinsic properties do not.  The distinction is important because the extrinsic terms act as simple convolutions and translations to the cloned spectrum.  We therefore define an augmented model, which we designate the ``extrinsic'' model, $\mathsf{S}_{\rm ext}$:

\begin{eqnarray}
    \mathsf{S}_{\rm ext}(\bm{\lambda}_z) = \mathsf{S}_{\rm clone}(\bm{\lambda}_\mathrm{z} - \frac{RV}{c}\lambda_\mathrm{c}) * \zeta \left(\frac{\bm{v}}{v\sin{i}}\right) \label{eqn_convolution}
\end{eqnarray}

@@

Most autodiff-aware \texttt{convolve} operators act in pixel-space, approximating kernels as numerically sampled functions.  There exists a special exponential spectral sample spacing that allows the convolution operators to work out-of-the-box for rotational broadening.  We therefore change the sampling from the native stellar and telluric wavelength coordinate grids, $\bm{\lambda}_S$ and $\bm{\lambda}_T$, to a new exponentially sampled wavelength grid, denoted with the subscript $Z$:

\begin{eqnarray}
    \bm{\lambda}_Z = \lambda_0  \exp{\frac{\bm{v}-v_0}{c}}
\end{eqnarray}

where $\bm{v}-v_0$ is the velocity vector going from zero to the velocity associated with the largest wavelength, with \emph{linear} spaced velocity samples.  We choose a sampling in velocity space of 0.5 km/s, which corresponds to about $10\times$ finer than the instrumental resolving power of HPF, and delivers a minimum and maximum wavelength spacing of 0.013 and 0.024 \AA$/\text{pixel}$ respectively for the HPF bandwidth.

Operationally, the radial velocity shift gets applied to the line center positions rather than scaling the entire wavelength grid point coordinates, $\bm{\lambda}_Z$. This approximate choice ignores relativistic effects, but yields a convenience: it cleanly makes the $RV$ autodiff-aware, meaning that an infinitesimal change to the RV value can be sensed through backpropagation by affecting only the line center positions.

@@

\begin{eqnarray}
    \mathsf{M}_{\rm joint} = \mathsf{S}_{\rm ext}(\bm{\lambda}_Z) \odot \mathsf{T}_{\rm clone}(\bm{\lambda}_Z)
\end{eqnarray}
